### 2020-07-31 Thesis/Source & Citations/Citations.tex
30 July 2020
- Cleaned up REPO
- Deleted excess files
- Added citations files

% ================================
% Landon Buell
% Kevin Short
% Physics 799%
% 30 July 2020 
% ================================

\documentclass[12pt,letterpaper]{article}
\usepackage[top=2.5cm,left=2.5cm,right=2.5cm]{geometry}

\begin{document}

% ================================================================

\begin{thebibliography}{9}
\bibliographystyle{apalike}

\bibitem{Geron}
Geron, Aurelien. \textit{Hands-on Machine Learning with Scikit-Learn and TensorFlow: Concepts, Tools, and Techniques to Build Intelligent Systems}. O'Reilly, 2017.

\bibitem{Geron2}
Geron, Aurelien. \textit{Hands-on Machine Learning with Scikit-Learn and TensorFlow: Concepts, Tools, and Techniques to Build Intelligent Systems}. 2nd ed., O'Reilly, 2019.

\bibitem{Goodfellow}
Goodfellow, Ian, et al.\textit{Deep Learning}. MIT Press, 2017.

\bibitem{James}
James, Gareth, et al. \textit{An Introduction to Statistical Learning with Applications in R}. Springer, 2017.

\bibitem{Kahn}
Khan, M. Kashif Saeed, and Wasfi G. Al-Khatib. “Machine-Learning Based Classification of Speech and Music.” Multimedia Systems, vol. 12, no. 1, 2006, pp. 55–67., doi:10.1007/s00530-006-0034-0.

\bibitem{Levine}
Levine, Daniel S. \textit{Introduction to Neural and Cognitive Modeling}. 3rd ed., Routledge, 2019.

\bibitem{Liu}
Liu, Zhu, et al. "Audio Feature Extraction and Analysis for Scene Segmentation and Classiﬁcation." Journal of VLSI Signal Processing, vol. 20, 1998, pp. 61–79.

\bibitem{Loy}
Loy, James , \textit{Neural Network Projects with Python}. Packt Publishing, 2019

\bibitem{McCulloch}
McCulloch, Warren S., and Walter Pitts. "A Logical Calculus of the Ideas Immanent in Nervous Activity." \textit{The Bulletin of Mathematical Biophysics}, vol. 5, no. 4, 1943, pp. 115–133.

\bibitem{Mierswa}
Mierswa, Ingo, and Katharina Morik. "Automatic Feature Extraction for Classifying Audio Data." \textit{Machine Learning}, vol. 58, no. 2-3, 2005, pp. 127–149., doi:10.1007/s10994-005-5824-7.

\bibitem{Peatross}
Peatross, Justin, and Michael Ware. \textit{Physics of Light and Optics.} Brigham Young University, Department of Physics, 2015.

\bibitem{Petrik}
Petrik, Marek. "Introduction to Deep Learning." Machine Learning. 20 April. 2020, Durham, New Hampshire.

\bibitem{Short}
Short, K. and Garcia R.A. 2006. "Signal Analysis Using the Complex Spectral Phase Evolution (CSPE) Method." AES: \textit{Audio Engineering Society Convention Paper}.

\bibitem{Virtanen}
Virtanen, Tuomas, et al. \textit{Computational Analysis of Sound Scenes and Events.} Springer, 2018.

\bibitem{Zhang}
Zhang, Tong, and C.-C. Jay Kuo. “Content-Based Classification and Retrieval of Audio.” \textit{Advanced Signal Processing Algorithms, Architectures, and Implementations VIII}, 2 Oct. 1998, pp. 432–443., doi:10.1117/12.325703.

\end{thebibliography}

% ================================================================

\end{document}